\textbf{Developmental and postural changes in children's visual access to faces}

### 2018-01-17 writeup-old/xsface.tex

---
title: "Developmental and postural changes in children's visual access to faces"
short-title: "Children's visual access to faces"
output: kmr::apa_manuscript
csl: apa6.csl
bibliography: xsface.bib

document-params: "a4paper,man,apacite,floatsintext,longtable"

bib-tex: "xsface.bib"

author-information:
    - \author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}

affiliation-information:
    - \affiliation{Department of Psychology, Stanford University}

author-note:
    "Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation. This research was supported by a John Merck Scholars grant to MCF. An earlier version of this work was presented to the Cognitive Science Society in @frank2013. Please address correspondence to Michael C. Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."

abstract: 
    "The faces of other people are a critical information source for young children. During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently. We used a head-mounted camera in conjunction with a face-detection system to explore the eff  ects of these changes on children's visual access to their caregivers' faces during an in-lab play session. In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture. These changes may translate into changes in the accessibility of social information during language learning. We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally."

keywords:
    "social cognition; face-perception; infancy; locomotion; head-cameras"
---

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
rm(list=ls())
library(readr)
# library(magrittr)
library(dplyr)
library(tidyr)
library(xtable)
# library(lubridate)
```

<!-- % A father offers his young daughter a novel object: a bright yellow feather duster. A few moments after she accepts the toy, he remarks, ``Isn't the zem funny?'' Her father may still be talking about the feather duster, or he may be describing a new object.  To find out she has access to a simple and reliable method: she can look to his face to infer the direction of his attention. -->

The ability to follow social signals like eye-gaze is an important part of early social cognition [@scaife1975;@tomasello2009]

In addition to being an important part of early social development, the ability to follow gaze is a strong predictor of children's early language development. For example, @brooks2005 found that children who followed an experimenter's gaze better before their first birthday had larger vocabularies at 18 months. Similarly, @carpenter1998 found that children's level of joint engagement (as well as the degree to which mothers followed the child's focus of attention in their labeling) predicted vocabulary growth in both language production and comprehension. These studies suggest that children's social environment plays a powerful supportive role in language learning.

But at the same time as children are beginning to learn their first words, their view of the world is changing radically [@adolph2007]. As speechless infants, they are unable to locomote independently. Before their first birthday, they begin crawling; soon after, they begin to walk independently. Infants' visual field is subject to the whims of their caregivers, but caregivers often place them in positions conducive to joint attention. In contrast, toddlers determine their own input to a much greater degree, but as a consequence they spend much of their time in a world primarily populated by knees. These postural and locomotor changes may have a profound effect on what children see.

A recent study suggests the possibility of links between motor milestones, social cognition, and language. @walle2014 noted robust correlations between children's ability to walk and their vocabulary, both receptive and productive. On the basis of an observational study of parent input, they speculated that the emergence of walking may change the ability of the child to access social information (because walking toddlers see more of the social world than crawling infants). Accessing more social information may in turn allow children to discover word meanings more effectively.

Recent methodological developments have the potential to provide data that would allow this hypothesis to be tested. The availability of head-mounted cameras and eye-trackers allows for the measurement of children's naturalistic environment in a way that was not previously possible. @yoshida2008 gave the first demonstration of the radical differences between toddler and adult perspectives on the social world, with toddlers' visual field being dominated by hands and objects much more than that of adults. More recent work has used head-mounted eye-tracking methods to measure young toddlers' fixations [@franchak2011], also finding that children look relatively infrequently at their mothers' faces in naturalistic play.

These methods are now being applied to understand inputs to language acquisition. Work by Yu, Smith, and colleagues suggests that word learning is facilitated when parents and children create moments in which the visual field is dominated by a single object [@smith2011;@yuinpress]. Some data even suggest that young children's restricted viewpoint may be more effective for learning words than the comparable adult perspective [@yurovsky2012]. Together, this body of evidence suggests that measuring infants' perspective---and how it changes in motor development---is a critical part of understanding early language learning.

@@

# Methods

## Participants

```{r echo=FALSE, include=FALSE}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
       function(x) {
         x <- as.numeric(x)
         x[1]+x[2]/60
       }
)

all_subs %>% 
  group_by(reason) %>%
  summarise(n = n()) %>% 
  kable

# # exclusions by age
# all_subs %>% 
#   group_by(reason, age_group) %>%
#   summarise(n = n()) %>%
#   spread(age_group, n) %>%
#   xtable %>%
#   print(digits = 2)

all_subs %>%
  left_join(inc_subs) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc), 
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE), 
            video_length = mean(len.min, na.rm=TRUE), 
            female = mean(gender=="female", na.rm=TRUE)) %>%
  xtable %>%
  print(digits = 2, include.rownames=FALSE)
```

@@

In order to compile this final sample, we tested a substantially larger group of children, who were excluded for the following reasons: 
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing CDI data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.

## Head-mounted camera

@@

Our head-mounted camera ("headcam") is composed of a small, inexpensive MD80 model camera attached to a soft elastic headband from a camping headlamp. An aftermarket fisheye lens intended for iPhones and other Apple devices is attached to increase view angle. The total cost of each camera is approximately \$60. The camera captures 720x480 pixel images at approx. 25 frames per second, and has battery life of 60--90 minutes. Without the fisheye lens, the viewing angle for the camera is $32^{\circ}$ horizontal by $24^{\circ}$ vertical; with the fisheye, $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left). The device is pictured in Figure \ref{fig:headcam}, right. Instructions for creating the camera are given in detail at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.

The vertical field of view of the camera was considerably smaller than the child's approximate vertical field of view, which---even at 6-7 months---spans around 100--120$^{\circ}$ in the vertical dimension [@mayer1988;@cummings1988]. We were therefore faced with a choice in the orientation of the camera. If we chose a lower or higher orientation, we would be at risk of truncating either the child's own hands and physically proximate objects, or the faces of the adults around the child. Yet if we chose the middle orientation, we would still be at risk of underestimating the proportion of faces viewed by the child. Thus, for the purposes of the current study---measuring visual access to faces---we chose to orient the camera towards the upper part of the visual field. While this orientation decreased our chances of recording the objects being manipulated by the child, it nevertheless allowed us to capture the majority of the faces in the child's visual field.

## Procedure

Before coming to the lab, caregivers completed the MacArthur-Bates Communicative Development Inventories Words and Gestures form [@fenson2007], reporting on the words that they believed that their child understood and produced.

@@

In the waiting room, the experimenter placed the headcam on children's heads after they had time to adjust to the environment. For children who resisted wearing the headcam, the experimenter used distractor techniques (presenting stimulating toys or engaging the children in hand-occupying activities) intended to keep children's focus elsewhere and prevent them from taking off the camera [@yoshida2008]. Once the child was wearing the camera comfortably for a period of time, child and caregiver were escorted to the playroom.

In the playroom, the experimenter presented the child's parent with a box containing three labeled pairs of objects, each consisting of a familiar and a novel object (e.g. a ball and a feather duster, marked as a "zem"). Parents confirmed that the child had not previously seen the novel toys. Parents were instructed to play with the object pairs with the child, one at a time, "as they typically would" and to use the novel labels to refer to the three toys. After giving these instructions, the experimenter left the room for a period of approximately 15 minutes. During this time, a tripod-mounted camera recorded video from a corner of the room and the headcam captured video from the child's perspective.

## Data Processing and Annotation

@@

All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded a substantial amount of video: a total of
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

### Posture and Orientation Annotation

One major goal of our study was to understand the relationship between children's posture and their access to information from the faces of their caregiver. To investigate this relationship, we created a set of annotations for the child's physical posture (e.g. standing, sitting) and orientation of the caregiver relative to the child (e.g. in front of, behind, close, far away). For each headcam video, a coder used OpenSHAPA software to annotate both orientation and posture [@adolph2012].

Orientation was initially categorized as having the caregiver in front, to the side, or behind the child, and close (defined informally as within arm's reach) or farther away. Because of data sparsity, we consolidated this scheme into three categories: close to the caregiver with the caregiver either in front or on the side, farther from the caregiver again with caregiver either in front or on the side, and a global category of caregiver behind the child. Posture was categorized as being held/carried, lying face-up, sitting, prone (crawling or lying), standing, or other. Data from when the child was out of view of the tripod camera was marked as uncodable and excluded from these annotations.
% A second coder coded XYZ videos; their categorizations were reliable at XYZ.

### Labeling Annotation

@@

## Face Detection

@@

### Face detection algorithms

The first algorithm was based on freely available computer vision tools [@bradski2008] and is described in depth in our previous work [@frank2012b]. This system had two parts. The first was the application of a set of four Haar-style face detection filters [@viola2004] to each frame of the videos independently. These detectors each provide information about whether a face is present in the frame as well as size and position for any detections. In a second step, these detections are then combined via a hidden Markov model (HMM), trained on hand-annotated data (see Appendix). The HMM model (which performed nearly as well as the more complex and computationally-intensive Conditional Random Field model used in our previous work) attempted to estimate whether a face was truly present in each frame of the videos, using as its input the number of Haar detectors that were active in any given frame.

The second algorithm that we evaluated was a semi-automated adaptive tracker-by-detection (SAATD). The algorithm required manual user input (selecting a single face example per video) for its initialization, but then needed no additional training data. The tracker is based on @kalal2010, which uses patches in the trajectory of an optical-flow based tracker [@lucas1981] to train and update a face detector. The optical flow tracker and the face detector work in parallel.
<!-- % For detection, we keep positive and negative templates of previously detected faces. -->
<!-- If the face detector finds a location in a new frame exhibiting a high similarity to its stored template, the tracker is re-initialised on that location. Otherwise, the tracker uses the optical flow to decide the location of a face in the new frame. -->
<!--  % The face detector is trained using visual features [@lepetit2005} and the detection is performed using a random ferns classifier [@ozuysal2007}. -->
<!-- The primary advantage of the SAATD algorithm is the use of motion for face detection: Following the movement of the pixels that define a face it is possible for the algorithm to adapt to new morphologies (i.e. different face poses). -->
<!-- %One disadvantage, however, is the necessity of manual initialization. -->

<!-- % [1] Z. Kalal, J. Matas, and K. Mikolajczyk. P-N learning: Bootstrapping binary classiﬁers by structural constraints. In 2010 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pages 49–56. IEEE, June 2010. -->

<!-- % [2] B. D. Lucas and T. Kanade. An iterative image registration technique with an application to stereo vision. In Proceedings of the International Joint Conference on Artiﬁcial Intelligence, pages 674–679, 1981 -->

<!-- % [3] V. Lepetit, P. Lagger, and P. Fua. Randomized trees for Real-Time keypoint recognition. -->
<!-- % In IEEE Computer Society Conference on Computer Vision and Pattern Recognition, volume 2, pages 775–781, Los Alamitos, CA, USA, 2005. IEEE Computer Society. -->

<!-- % [4] M. Ozuysal, P. Fua, and V. Lepetit. Fast keypoint recognition in ten lines of code. In IEEE Conference on Computer Vision and Pattern Recognition, Los Alamitos, CA, USA, June 2007. IEEE. -->

### Detector evaluation

@@

To ensure that our evaluation was not biased by the relatively rare appearance of faces in the dataset, we annotated two samples, both a random sample from the data and a sample with a high-density of faces. We selected 1 minute of interaction for each age group, divided evenly across the four dyads at that age. For each dyad, we divided the recorded video into contiguous 1 s segments and selected 16 in accordance with two criteria. First, 8 of these segments were selected by choosing the parts of the videos highest in face detection (*high density sample*). To be fair to both algorithms, half of this was chosen from the segments with the most HMM detections and half were chosen from the segments with the most SAATD detections. The remaining segments were chosen by randomly sampling from segments not yet selected for coding (*random sample*).

@@

# Results

@@

## Changes in Posture and Orientation

<!-- \begin{figure}[t] -->
<!-- \centering -->
<!-- \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- \caption{\label{fig:posture} Proportion time in each posture, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
<!-- \end{figure} -->

<!-- \begin{figure}[t] -->
<!-- \centering -->
<!-- \includegraphics[width=5in]{figures/orientation.pdf} -->
<!-- \caption{\label{fig:orientation} Proportion time in each orientation, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
<!-- \end{figure} -->

@@

## Access to Faces

<!-- % \begin{figure}[ -->
<!-- % \begin{figure}[t] -->
<!-- % \centering -->
<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel). We omit the lying face-up posture due to data sparsity. Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
<!-- % \end{figure}t] -->
<!-- % \centering -->
<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel). We omit the lying face-up posture due to data sparsity. Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
<!-- % \end{figure} -->

@@

a
## Access to Faces During Labeling

Our final analysis concerned the accessibility of caregivers' faces during labeling events. @franchak2011 found that referential speech was marginally more likely to draw toddlers' attention to mothers' faces. We were similarly interested in whether looking at faces occurred during labeling. Accordingly, we used the labeling annotations for each child to identify the 2s before and after each labeling event. We then computed the proportion face detections within this window across ages.

@@

<!-- % \begin{figure} -->
<!-- % \begin{center} -->
<!-- %   \includegraphics[width=3in]{figures/naming_faces_diff.pdf} -->
<!-- % \end{center} -->
<!-- % \caption{\label{fig:naming_face} Proportion of faces detected in a 4s window of time centered around labeling events, plotted by age group and whether the word was familiar or novel. Error bars show 95\% confidence intervals. 4-month-olds are omitted due to the limited number of total face detections for this group.} -->
<!-- % % \vspace{-ex} -->
<!-- % \end{figure} -->

# General Discussion

@@

The measures developed here have broad applicability to the study of individual and cultural differences. Since the physical circumstances of child rearing vary widely across households and across cultures, there may be important and predictable differences in children's visual experience. As suggested by the correlations between walking and vocabulary development [@walleunderreview], postural development may have substantial downstream consequences for language. For example, shifts in how infants are placed in particular postures by strollers or carriers [@zeedyk2008] or how their motor development is encouraged by parent practices [@bril1986] may lead to differences in social input which in turn affect their language learning. Since our variant of the headcam method is both inexpensive and highly portable, we have been able to deploy it in children's homes with some success; it may thus be a valuable tool for investigating differences in child-rearing practices.

A deep body of work uses children's linguistic input---measured using audio recordings---to understand the learning mechanisms underlying vocabulary acquisition [@huttenlocher1991;@hart1995;@fernald2006]. There have been some important initial successes in using visual input to predict language uptake [@yuinpress]. Nevertheless, we have a long way to go before our knowledge about children's visual input parallels our understanding of their linguistic environment. Coming to such an understanding will require the creation of both corpus resources and automated tools such as those we have begun to develop here.

# View angle

We note that previous studies have shown that children's head movements in the horizontal dimension are approximated by (though are slightly lagged by) their head movements [@yoshida2008]. Our own experience with the current apparatus ratifies these conclusions for the horizontal field but suggests that head movements in the vertical field are less reliable. Hence, these studies may run the risk of underestimating the proportion of faces actually seen by children.

# References

>>>>

% Template for APA submission with R Markdown

% Stuff changed from PLOS Template
\documentclass[a4paper,man,apacite,floatsintext,longtable]{apa6}
\usepackage{apacite}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

%\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

% ALL OF THE TITLE PAGE INFORMATION IS SPECIFIED IN THE YAML
\title{\textbf{Developmental and postural changes in children's visual access to faces}}
\shorttitle{Children's visual access to faces}

\author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}

\affiliation{Department of Psychology, Stanford University}

\authornote{Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the
Language and Cognition Lab for help in recruitment, data collection, and
annotation. This research was supported by a John Merck Scholars grant
to MCF. An earlier version of this work was presented to the Cognitive
Science Society in Michael C Frank, Simmons, Yurovsky, \& Pusiol (2013).
Please address correspondence to Michael C. Frank, Department of
Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford,
CA, 94305, tel: (650) 724-4003, email:
\href{mailto:exttt\%7Bmcfrank@stanford.edu\%7D}{\nolinkurl{exttt\{mcfrank@stanford.edu\}}}.}
\abstract{The faces of other people are a critical information source for young
children. During early development, children undergo significant
postural and locomotor development, changing from lying and sitting
infants to toddlers who walk independently. We used a head-mounted
camera in conjunction with a face-detection system to explore the eff
ects of these changes on children's visual access to their caregivers'
faces during an in-lab play session. In a cross-sectional sample of
8--16 month old children, we found substantial changes in face
accessibility based on age and posture. These changes may translate into
changes in the accessibility of social information during language
learning. We make our corpus available for reanalysis, and discuss
strengths and limitations of the headcam method more generally.}
\keywords{social cognition; face-perception; infancy; locomotion; head-cameras}

\begin{document}
\maketitle

The ability to follow social signals like eye-gaze is an important part
of early social cognition (Scaife \& Bruner, 1975; M. Tomasello, 2009)

In addition to being an important part of early social development, the
ability to follow gaze is a strong predictor of children's early
language development. For example, Brooks \& Meltzoff (2005) found that
children who followed an experimenter's gaze better before their first
birthday had larger vocabularies at 18 months. Similarly, Carpenter,
Nagell, \& Tomasello (1998) found that children's level of joint
engagement (as well as the degree to which mothers followed the child's
focus of attention in their labeling) predicted vocabulary growth in
both language production and comprehension. These studies suggest that
children's social environment plays a powerful supportive role in
language learning.

But at the same time as children are beginning to learn their first
words, their view of the world is changing radically (Adolph \& Berger,
2007). As speechless infants, they are unable to locomote independently.
Before their first birthday, they begin crawling; soon after, they begin
to walk independently. Infants' visual field is subject to the whims of
their caregivers, but caregivers often place them in positions conducive
to joint attention. In contrast, toddlers determine their own input to a
much greater degree, but as a consequence they spend much of their time
in a world primarily populated by knees. These postural and locomotor
changes may have a profound effect on what children see.

A recent study suggests the possibility of links between motor
milestones, social cognition, and language. Walle \& Campos (2014) noted
robust correlations between children's ability to walk and their
vocabulary, both receptive and productive. On the basis of an
observational study of parent input, they speculated that the emergence
of walking may change the ability of the child to access social
information (because walking toddlers see more of the social world than
crawling infants). Accessing more social information may in turn allow
children to discover word meanings more effectively.

Recent methodological developments have the potential to provide data
that would allow this hypothesis to be tested. The availability of
head-mounted cameras and eye-trackers allows for the measurement of
children's naturalistic environment in a way that was not previously
possible. Yoshida \& Smith (2008) gave the first demonstration of the
radical differences between toddler and adult perspectives on the social
world, with toddlers' visual field being dominated by hands and objects
much more than that of adults. More recent work has used head-mounted
eye-tracking methods to measure young toddlers' fixations (Franchak,
Kretch, Soska, \& Adolph, 2011), also finding that children look
relatively infrequently at their mothers' faces in naturalistic play.

These methods are now being applied to understand inputs to language
acquisition. Work by Yu, Smith, and colleagues suggests that word
learning is facilitated when parents and children create moments in
which the visual field is dominated by a single object (L. B. Smith, Yu,
\& Pereira, 2011; Yu \& Smith, in press). Some data even suggest that
young children's restricted viewpoint may be more effective for learning
words than the comparable adult perspective (D. Yurovsky, Smith, \& Yu,
in press). Together, this body of evidence suggests that measuring
infants' perspective---and how it changes in motor development---is a
critical part of understanding early language learning.

@@

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

@@

In order to compile this final sample, we tested a substantially larger
group of children, who were excluded for the following reasons: 20 for
technical issues related to the headcam, 15 for failing to wear the
headcam, 10 for fewer than 4 minutes of headcam footage, 5 for having
multiple adults present, 5 for missing CDI data, 2 for missing scene
camera footage, 1 for fussiness, and one excluded for sample symmetry.
All inclusion decisions were made independent of the results of
subsequent analyses.

\subsection{Head-mounted camera}\label{head-mounted-camera}

@@

Our head-mounted camera (``headcam'') is composed of a small,
inexpensive MD80 model camera attached to a soft elastic headband from a
camping headlamp. An aftermarket fisheye lens intended for iPhones and
other Apple devices is attached to increase view angle. The total cost
of each camera is approximately \$60. The camera captures 720x480 pixel
images at approx. 25 frames per second, and has battery life of 60--90
minutes. Without the fisheye lens, the viewing angle for the camera is
\(32^{\circ}\) horizontal by \(24^{\circ}\) vertical; with the fisheye,
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, left). The device is pictured in Figure
\ref{fig:headcam}, right. Instructions for creating the camera are given
in detail at
\url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.

The vertical field of view of the camera was considerably smaller than
the child's approximate vertical field of view, which---even at 6-7
months---spans around 100--120\(^{\circ}\) in the vertical dimension
(Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton, 1988; Mayer,
Fulton, \& Cummings, 1988). We were therefore faced with a choice in the
orientation of the camera. If we chose a lower or higher orientation, we
would be at risk of truncating either the child's own hands and
physically proximate objects, or the faces of the adults around the
child. Yet if we chose the middle orientation, we would still be at risk
of underestimating the proportion of faces viewed by the child. Thus,
for the purposes of the current study---measuring visual access to
faces---we chose to orient the camera towards the upper part of the
visual field. While this orientation decreased our chances of recording
the objects being manipulated by the child, it nevertheless allowed us
to capture the majority of the faces in the child's visual field.

\subsection{Procedure}\label{procedure}

Before coming to the lab, caregivers completed the MacArthur-Bates
Communicative Development Inventories Words and Gestures form (Fenson et
al., 2007), reporting on the words that they believed that their child
understood and produced.

@@

In the waiting room, the experimenter placed the headcam on children's
heads after they had time to adjust to the environment. For children who
resisted wearing the headcam, the experimenter used distractor
techniques (presenting stimulating toys or engaging the children in
hand-occupying activities) intended to keep children's focus elsewhere
and prevent them from taking off the camera (Yoshida \& Smith, 2008).
Once the child was wearing the camera comfortably for a period of time,
child and caregiver were escorted to the playroom.

In the playroom, the experimenter presented the child's parent with a
box containing three labeled pairs of objects, each consisting of a
familiar and a novel object (e.g.~a ball and a feather duster, marked as
a ``zem''). Parents confirmed that the child had not previously seen the
novel toys. Parents were instructed to play with the object pairs with
the child, one at a time, ``as they typically would'' and to use the
novel labels to refer to the three toys. After giving these
instructions, the experimenter left the room for a period of
approximately 15 minutes. During this time, a tripod-mounted camera
recorded video from a corner of the room and the headcam captured video
from the child's perspective.

\subsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

@@

All headcam videos were cropped to exclude the period of entry to the
playroom and were automatically synchronized with the tripod-mounted
videos using FinalCut Pro Software. These sessions yielded a substantial
amount of video: a total of 515 minutes (almost a milion frames), with
an average video length of 8.58 minutes (min = 4.53, max = 19.35).

\subsubsection{Posture and Orientation
Annotation}\label{posture-and-orientation-annotation}

One major goal of our study was to understand the relationship between
children's posture and their access to information from the faces of
their caregiver. To investigate this relationship, we created a set of
annotations for the child's physical posture (e.g.~standing, sitting)
and orientation of the caregiver relative to the child (e.g.~in front
of, behind, close, far away). For each headcam video, a coder used
OpenSHAPA software to annotate both orientation and posture (Adolph,
Gilmore, Freeman, Sanderson, \& Millman, 2012).

Orientation was initially categorized as having the caregiver in front,
to the side, or behind the child, and close (defined informally as
within arm's reach) or farther away. Because of data sparsity, we
consolidated this scheme into three categories: close to the caregiver
with the caregiver either in front or on the side, farther from the
caregiver again with caregiver either in front or on the side, and a
global category of caregiver behind the child. Posture was categorized
as being held/carried, lying face-up, sitting, prone (crawling or
lying), standing, or other. Data from when the child was out of view of
the tripod camera was marked as uncodable and excluded from these
annotations. \% A second coder coded XYZ videos; their categorizations
were reliable at XYZ.

\subsubsection{Labeling Annotation}\label{labeling-annotation}

@@

\subsection{Face Detection}\label{face-detection}

@@

\subsubsection{Face detection
algorithms}\label{face-detection-algorithms}

The first algorithm was based on freely available computer vision tools
(Bradski \& Kaehler, 2008) and is described in depth in our previous
work (M. C. Frank, 2012). This system had two parts. The first was the
application of a set of four Haar-style face detection filters (Viola \&
Jones, 2004) to each frame of the videos independently. These detectors
each provide information about whether a face is present in the frame as
well as size and position for any detections. In a second step, these
detections are then combined via a hidden Markov model (HMM), trained on
hand-annotated data (see Appendix). The HMM model (which performed
nearly as well as the more complex and computationally-intensive
Conditional Random Field model used in our previous work) attempted to
estimate whether a face was truly present in each frame of the videos,
using as its input the number of Haar detectors that were active in any
given frame.

The second algorithm that we evaluated was a semi-automated adaptive
tracker-by-detection (SAATD). The algorithm required manual user input
(selecting a single face example per video) for its initialization, but
then needed no additional training data. The tracker is based on Kalal,
Mikolajczyk, \& Matas (2010), which uses patches in the trajectory of an
optical-flow based tracker (Lucas \& Kanade, 1981) to train and update a
face detector. The optical flow tracker and the face detector work in
parallel.

\subsubsection{Detector evaluation}\label{detector-evaluation}

@@

To ensure that our evaluation was not biased by the relatively rare
appearance of faces in the dataset, we annotated two samples, both a
random sample from the data and a sample with a high-density of faces.
We selected 1 minute of interaction for each age group, divided evenly
across the four dyads at that age. For each dyad, we divided the
recorded video into contiguous 1 s segments and selected 16 in
accordance with two criteria. First, 8 of these segments were selected
by choosing the parts of the videos highest in face detection
(\emph{high density sample}). To be fair to both algorithms, half of
this was chosen from the segments with the most HMM detections and half
were chosen from the segments with the most SAATD detections. The
remaining segments were chosen by randomly sampling from segments not
yet selected for coding (\emph{random sample}).

@@

\section{Results}\label{results}

@@

\subsection{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

@@

\subsection{Access to Faces}\label{access-to-faces}

@@

Across ages, the carrying and prone postures resulted in the smallest
number of faces seen, while standing and sitting resulted in far more.
These postures both presented opportunities for seeing faces in large
part because parents were sitting or lying on the floor with children.
Although far fewer faces were seen when the caregiver was behind the
child,\footnote{Since orientation was coded via body posture, faces seen
  while the caregiver was behind the child were due to children looking
  over their shoulder.} both the close and far positions resulted in
approximately equal proportions of face detections.

\subsection{Access to Faces During
Labeling}\label{access-to-faces-during-labeling}

Our final analysis concerned the accessibility of caregivers' faces
during labeling events. Franchak et al. (2011) found that referential
speech was marginally more likely to draw toddlers' attention to
mothers' faces. We were similarly interested in whether looking at faces
occurred during labeling. Accordingly, we used the labeling annotations
for each child to identify the 2s before and after each labeling event.
We then computed the proportion face detections within this window
across ages.

@@

\section{General Discussion}\label{general-discussion}

@@

The measures developed here have broad applicability to the study of
individual and cultural differences. Since the physical circumstances of
child rearing vary widely across households and across cultures, there
may be important and predictable differences in children's visual
experience. As suggested by the correlations between walking and
vocabulary development ({\textbf{???}}), postural development may have
substantial downstream consequences for language. For example, shifts in
how infants are placed in particular postures by strollers or carriers
(Zeedyk, 2008) or how their motor development is encouraged by parent
practices (Bril \& Sabatier, 1986) may lead to differences in social
input which in turn affect their language learning. Since our variant of
the headcam method is both inexpensive and highly portable, we have been
able to deploy it in children's homes with some success; it may thus be
a valuable tool for investigating differences in child-rearing
practices.

A deep body of work uses children's linguistic input---measured using
audio recordings---to understand the learning mechanisms underlying
vocabulary acquisition (Fernald, Perfors, \& Marchman, 2006; Hart \&
Risley, 1995; Huttenlocher, Haight, Bryk, Seltzer, \& Lyons, 1991).
There have been some important initial successes in using visual input
to predict language uptake (Yu \& Smith, in press). Nevertheless, we
have a long way to go before our knowledge about children's visual input
parallels our understanding of their linguistic environment. Coming to
such an understanding will require the creation of both corpus resources
and automated tools such as those we have begun to develop here.

\section{View angle}\label{view-angle}

We note that previous studies have shown that children's head movements
in the horizontal dimension are approximated by (though are slightly
lagged by) their head movements (Yoshida \& Smith, 2008). Our own
experience with the current apparatus ratifies these conclusions for the
horizontal field but suggests that head movements in the vertical field
are less reliable. Hence, these studies may run the risk of
underestimating the proportion of faces actually seen by children.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-adolph2007}{}
Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
of child psychology}. Wiley Online Library.

\hypertarget{ref-adolph2012}{}
Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
(2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-bradski2008}{}
Bradski, G., \& Kaehler, A. (2008). \emph{Learning OpenCV: Computer
vision with the OpenCV library}. O'Reilly Media.

\hypertarget{ref-bril1986}{}
Bril, B., \& Sabatier, C. (1986). The cultural context of motor
development: Postural manipulations in the daily life of Bambara babies
(Mali). \emph{International Journal of Behavioral Development},
\emph{9}(4), 439--453.

\hypertarget{ref-brooks2005}{}
Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
and its relation to language. \emph{Developmental Science}, \emph{8}(6),
535--543.

\hypertarget{ref-carpenter1998}{}
Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
joint attention, and communicative competence from 9 to 15 months of
age. \emph{Monographs of the Society for Research in Child Development},
\emph{63}(4).

\hypertarget{ref-cummings1988}{}
Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
(1988). Visual fields of young children. \emph{Behavioural and Brain
Research}, \emph{29}(1), 7--16.
